Theories of deliberated choice

### 2024-05-03 simple.tex
Conj closing

  Define the extension of a set of attitudes $S \wsub P^{\allhists}$ to its observably included attitudes as the set of attitudes $R(S) = \set{g \in P^{\allhists} \suchthat \forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}}$.
  Note that $S \wsub R(S)$. 
  Also note that $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$.
\end{definition}
\begin{theorem}
  \label{th:RR}
  $\forall S \wsub P^{\allhists}: R(R(S)) \wsub R(S)$.
\end{theorem}
\begin{proof}
  Consider any $g \in R(R(S))$.
  Consider any $H \in \finite(\incr(\allhists))$.
  There exists $f \in R(S)$ such that $\restr{g}{H} = \restr{f}{H}$.
  There exists $f' \in S$ such that $\restr{f}{H} = \restr{f'}{H}$.
  Thus, $\restr{g}{H} = \restr{f'}{H}$.
  Thus, $g \in R(S)$.
\end{proof}
Define the closure under observable inclusion of a claim about types $S$ as $R^\infty(S)$, equivalently, as the smallest set $S'$ such that $S \wsub S'$ and $R(S') \wsub S'$, equivalently, as indicated by \cref{th:RR}, as $R(S)$.
\begin{remark}
  A few more, maybe not useful, definitions.

  A function $l: \allhists \to \powerset{P^{\allhists}}$ is decreasing iff $\forall \hist \wsub \histp: l(\histp) \wsub l(\hist)$.
  Let $S \wsub P^{\allhists}$ be a set of attitudes.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-point-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_\hist$.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-sequence-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_{\intvl{\hist}}$.
\end{remark}
Let $\intvl{\hist} = (\hist_i)_{i \in \intvl{0, \card{\hist}}} \in \finite(\incr(\allhists))$ denote the increasing sequence of paths from $\emptyset$ to $\hist$.
\begin{theorem}
  The $S$-labeling function associated to an attitude $f \in P^{\allhists}$ is the function $l^S_f: \allhists \to \powerset{P^{\allhists}}$ defined as $l^S_f(\hist) = \set{g \in S \suchthat g(\hist) = f(\hist)}$.

  An attitude $f$ is observably included in a set of attitudes $S$ iff $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
\end{theorem}
\begin{proof}
  Consider $f$ with $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
  Consider any $H \in \finite(\incr(\allhists))$.
  We know that $\cap_{\histp \wsub \hist}{l^S_f(\histp)} \neq \emptyset$, thus in particular, defining $\hist = \max H$ and observing that $H \wsub \set{\histp \suchthat \histp \wsub \hist}$, we have that $\cap_{\histp \in H} l^S_f(\histp) \neq \emptyset$.
  Pick any $g \in \cap_{\histp \in H} l^S_f(\histp)$. Thus, $g \in S$.
  Also, $\forall \hist \in H: g \in l^S_f(\hist)$, thus,
  $\forall \hist \in H: g(\hist) = f(\hist)$.

@@

The following result follows from \cref{th:typesfalsifiable}.
\begin{theorem}
  If a set of attitudes $S \wsub P^{\allhists}$ is not closed under observable inclusion, then it is not falsifiable and not trivial.
\end{theorem}
\begin{theorem}
  Given any situation, if the claim associated to the set of attitudes $R^\infty({\allleadsto}(I))$ is trivial, then there exists no claim associated to a set of attitudes that holds and is falsifiable.
\end{theorem}
\begin{proof}
  Consider a claim related to a set of attitudes $S$ that holds and that is falsifiable or trivial.
  For it to hold, we need ${\allleadsto}(I) \wsub S$.
  From \cref{th:typesfalsifiable}, for it to be falsifiable or trivial, we need that $\forall g \in P^{\allhists} \setminus S: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in S: \restr{g}{H} \neq \restr{f}{H}$. In other words, all $g$ must be observably excludable from $S$. In other words, $S$ must be closed under observable inclusion. Thus, we need $R^\infty(S) \wsub S$.
  Also, because $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$, we need
  $R^\infty({\allleadsto}(I)) \wsub R^\infty(S) \wsub S$.
  It follows that $S$ is trivial.
\end{proof}
\begin{theorem}
  Given any situation, the claim associated to the set of attitudes $R^\infty({\allleadsto}(I))$ holds, is anonymous, and is falsifiable or trivial.
\end{theorem}

@@

  For example, $\allargs = \set{\ar_1, \ar_2}$ and ${\ileadstog}$ is $\phi$ when the path starts with $\ar_1$ (or is empty) and is $¬\phi$ when the path starts with $\ar_2$ and the attitude contains only people constant to $\phi$ (except that they map $\emptyset$ to anything) and people constant to $¬\phi$ (except that they map $\emptyset$ to anything), say, $(\phi, \phi)$, $(¬\phi, \phi)$, $(\phi, ¬\phi)$, $(¬\phi, ¬\phi)$.

>>>>

  Define the extension of a set of attitudes $S \wsub P^{\allhists}$ to its observably included attitudes as the set of attitudes $R(S) = \set{g \in P^{\allhists} \suchthat \forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}}$.
  Note that $S \wsub R(S)$. 
  Also note that $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$.
\end{definition}
\begin{theorem}
  \label{th:RR}
  $\forall S \wsub P^{\allhists}: R(R(S)) \wsub R(S)$.
\end{theorem}
\begin{proof}
  Consider any $g \in R(R(S))$.
  Consider any $H \in \finite(\incr(\allhists))$.
  There exists $f \in R(S)$ such that $\restr{g}{H} = \restr{f}{H}$.
  There exists $f' \in S$ such that $\restr{f}{H} = \restr{f'}{H}$.
  Thus, $\restr{g}{H} = \restr{f'}{H}$.
  Thus, $g \in R(S)$.
\end{proof}
Define the closure under observable inclusion of a claim about types $S$ as $R^\infty(S)$, equivalently, as the smallest set $S'$ such that $S \wsub S' = R(S')$, equivalently, as indicated by \cref{th:RR}, as $R(S)$.

Let $\intvl{\hist} = (\hist_i)_{i \in \intvl{0, \card{\hist}}} \in \finite(\incr(\allhists))$ denote the increasing sequence of paths from $\emptyset$ to $\hist$.
\begin{theorem}
  \label{th:caractR}
  The $S$-labeling function associated to an attitude $f \in P^{\allhists}$ is the function $l^S_f: \allhists \to \powerset{P^{\allhists}}$ defined as $l^S_f(\hist) = \set{g \in S \suchthat g(\hist) = f(\hist)}$.

  We have: $R(S) = \set{f \suchthat \forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset}$.
\end{theorem}
\begin{proof}
  Consider $f$ with $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
  Consider any $H \in \finite(\incr(\allhists))$.
  We know that $\cap_{\histp \wsub \hist}{l^S_f(\histp)} \neq \emptyset$, thus in particular, defining $\hist = \max H$ and observing that $H \wsub \set{\histp \suchthat \histp \wsub \hist}$, we have that $\cap_{\histp \in H} l^S_f(\histp) \neq \emptyset$.
  Pick any $g \in \cap_{\histp \in H} l^S_f(\histp)$. Thus, $g \in S$.
  Also, $\forall \hist \in H: g \in l^S_f(\hist)$, thus,
  $\forall \hist \in H: g(\hist) = f(\hist)$.

@@

The following result follows from \cref{th:typesfalsifiable}.
\begin{theorem}
  If a set of attitudes $S \wsub P^{\allhists}$ is not closed under observable inclusion, then it is not falsifiable and not trivial.
\end{theorem}
\begin{theorem}
  Given any situation, if the claim associated to the set of attitudes $R({\allleadsto}(I))$ is trivial, then there exists no claim associated to a set of attitudes that holds and is falsifiable.
\end{theorem}
\begin{proof}
  Consider a claim related to a set of attitudes $S$ that holds and that is falsifiable or trivial.
  For it to hold, we need ${\allleadsto}(I) \wsub S$.
  From \cref{th:typesfalsifiable}, for it to be falsifiable or trivial, we need that $\forall g \in P^{\allhists} \setminus S: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in S: \restr{g}{H} \neq \restr{f}{H}$. In other words, all $g$ must be observably excludable from $S$. In other words, $S$ must be closed under observable inclusion. Thus, we need $R(S) \wsub S$.
  Also, because $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$, we need
  $R({\allleadsto}(I)) \wsub R(S) \wsub S$.
  It follows that $S$ is trivial.
\end{proof}
\begin{theorem}
  Given any situation, the claim associated to the set of attitudes $R({\allleadsto}(I))$ holds, is anonymous, and is falsifiable or trivial.
\end{theorem}
\begin{example}
  Define $f^{(p, p')} \in P^{\allhists}$ as follows: $f^{(p, p')}(\emptyset) = p$ and $\forall \hist \in \allhists: \hist ≠ \emptyset \implies f^{(p, p')}(\hist) = p'$.

  Consider a situation such that ${\allleadsto}(I) = \set{f^{(p, p')} \suchthat p, p' \in \set{\phi, ¬\phi}}$, thus, containing only people constant to $\phi$ and people constant to $¬\phi$, except that they map $\emptyset$ to anything in $\set{\phi, ¬\phi}$.
  Define $F^p = \set{f^{(p, p')} \suchthat p' \in \set{\phi, ¬\phi}}$.

  \Cref{th:caractR} indicates that $R({\allleadsto}(I)) = \set{f \suchthat l^S_f(\emptyset) \in \set{F^\phi, F^{¬\phi}} \land \forall \ar \in \allargs: [\forall \hist \suchthat \hist_1 = \ar: l^S_f(\hist) \in l^S_f(\emptyset)] \land [\forall \hist, \histp \suchthat \hist_1 = \histp_1 = \ar: l^S_f(\hist) = l^S_f(\histp)]}$.

  Thus, $R({\allleadsto}(I))$ contains, for each mapping $m: \allargs \rightarrow \set{\phi, ¬\phi}$, the two functions $f_m^\phi, f_m^{¬\phi}$ defined as $f_m^\phi(\emptyset) = \phi$ and $\forall \ar \in \allargs: f_m^\phi((\ar, …)) = m(\ar)$, and similarly for $f_m^{¬\phi}$.

  The claim associated to $R({\allleadsto}(I))$ is not trivial: $\forall f \suchthat [\exists \hist, \histp ≠ \emptyset \suchthat \hist_1 = \histp_1 \land f(\hist) ≠ f(\histp)]: f \notin R({\allleadsto}(I))$.
\end{example}
\begin{example}
  Consider $\allargs = \set{\ar_1, \ar_2}$ and ${\allleadsto}(I) = \set{f \suchthat f(\ar_1) ≠ \phi \lor f(\ar_2) ≠ \phi}$, thus, $f \notin {\allleadsto}(I)$ iff $f(\ar_1) = f(\ar_2) = \phi$.

  Consider any $f \notin {\allleadsto}(I)$.
  We have $f \in R({\allleadsto}(I))$.
  Indeed, define $g_1 \in P^{\allhists}$ as $g_1 = f$ except that $g_1(\ar_2) = ¬\phi$, and $g_2 \in P^{\allhists}$ as $g_2 = f$ except that $g_2(\ar_1) = ¬\phi$.
  Then, $g_1, g_2 \in {\allleadsto}(I)$.
  Now consider any $H \in \finite(\incr(\allhists))$.
  If $(\ar_2) \notin H$, then $\restr{f}{H} = \restr{g_1}{H}$; otherwise, $\restr{f}{H} = \restr{g_2}{H}$.

  It follows that the claim associated to $R({\allleadsto}(I))$ is trivial.
\end{example}
\begin{remark}
  A few more, maybe not useful, definitions.

  A function $l: \allhists \to \powerset{P^{\allhists}}$ is decreasing iff $\forall \hist \wsub \histp: l(\histp) \wsub l(\hist)$.
  Let $S \wsub P^{\allhists}$ be a set of attitudes.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-point-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_\hist$.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-sequence-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_{\intvl{\hist}}$.
\end{remark}
\begin{conjecture}
  Given a non empty set of attitudes $S \wsub P^{\allhists}$, the claim associated to $R(S)$ is not trivial iff [there exists some $H \in \finite(\incr(\allhists))$ such that some non empty equivalence class $S_H \in S / =_H$ on $S$ over $H$ does not permit all continuations, meaning that $\exists \ar, p \suchthat \forall f \in S_H: f(\max H, \ar) ≠ p$]; or $\exists p \suchthat \forall f \in S: f(\emptyset) ≠ p$.
\end{conjecture}
\begin{proof}
  We want to show that $\exists f \in P^{\allhists}, \hist \in \allhists \suchthat \cap_{\histp \wsub \hist} l^S_f(\histp) = \emptyset$ is equivalent to the stated consequent.

  Assuming that the consequent holds, consider any $f$ such that $\restr{f}{H} = \restr{g}{H}$ for some (equivalently, for all) $g \in S$ and such that $f(\max H, \ar) = p$ (or pick $H = \ar = \emptyset$).
  We have $\cap_{\histp \wsub (\max H, \ar)} l^S_f(\histp) = \emptyset$.

  Consider any $f, \hist \suchthat \cap_{\histp \wsub \hist} l^S_f(\histp) = \emptyset$ and let’s prove that the consequent holds.
  As $S ≠ \emptyset$, $l^S_f(\hist) ≠ \emptyset$.
  Consider the longest subsequence of $\hist$ such that $l^S_f(\histp) ≠ \emptyset$.
\end{proof}

Generally,
define $S^p = \set{f \in P^{\allhists} \suchthat f(\emptyset) = p}$.
Then \cref{th:caractR} indicates that $R({\allleadsto}(I)) = \set{f \suchthat l^S_f(\emptyset) \in \set{S^p \suchthat p \in P} \land \forall \ar \in \allargs: [\forall \hist \suchthat \hist_1 = \ar: l^S_f(\hist) \in l^S_f(\emptyset)] \land [\forall \hist, \histp \suchthat \hist_1 = \histp_1 = \ar: l^S_f(\hist) = l^S_f(\histp)]}$.
TODO correct this.

@@

  For example,  and ${\ileadstog}$ is $\phi$ when the path starts with $\ar_1$ (or is empty) and is $¬\phi$ when the path starts with $\ar_2$ and
